---
title: "S3rd_visualizationfitSCcurves_SA12sumSCinvnode_CuiBP"
output: html_document
date: "2024-11-03"
---

This script is to generate fitted values from gam models. The predicted SC strength will be generated as the age was sampled from 6 to 22, with a total of 1000 data points, covariables will be set as median or mode.The data will be used to draw developmental trajectories. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(R.matlab)
library(mgcv)
library(psych)
library(tidyverse)
library(parallel)
library(scales)
library(openxlsx)
library(gratia)
library(ggplot2)
library(RColorBrewer)
library(paletteer)
library(patchwork)
rm(list = ls())

interfileFolder <- 'D:/xuxiaoyu/DMRI_network_development/SC_development/interdataFolder_ChineseCohort'
FigureFolder<-'D:/xuxiaoyu/DMRI_network_development/SC_development/Figure_ChineseCohort_final/SA12'
functionFolder <- 'D:/xuxiaoyu/DMRI_network_development/SC_development/Rcode_SCdevelopment/gamfunction'
resultFolder<-'D:/xuxiaoyu/DMRI_network_development/SC_development/results_ChineseCohort'

```

## 1. load data & function

```{r load_data}
#### load data
CVthr = 75
gamresultsum.SAorder.delLM<-readRDS(paste0(interfileFolder, '/gamresults78_sumSCinvnode_CV', CVthr,'_scale_TRUE.rds'))
gammodelsum<-readRDS(paste0(interfileFolder, '/gammodel78_sumSCinvnode_CV', CVthr,'_scale_TRUE.rds'))
SCdata <- readRDS(paste0(interfileFolder, '/SCdata_SA12_CV', CVthr,'_sumSCinvnode.sum.msmtcsd.combatAge.rds'))
derivative <- readRDS(paste0(resultFolder, '/derivative.df78_CV', CVthr,'.rds'))
#### source function
source(paste0(functionFolder, '/plotdata_generate.R'))
source(paste0(functionFolder, '/colorbarvalue.R'))
#### generate fitted values for developmental trajectories
agevector <- SCdata$Age
# set cluster
num_cores <- detectCores() - 8
cl <- makeCluster(num_cores)
clusterExport(cl, varlist = ls(), envir = .GlobalEnv)
invisible(clusterEvalQ(cl, {
  library(tidyverse)
  library(R.matlab)
  library(psych)
  library(gratia)
  library(mgcv)
  library(parallel)
  library(ggplot2)
  library(RColorBrewer)
  library(reshape)
  source(paste0(functionFolder, '/plotdata_generate.R'))
  source(paste0(functionFolder, '/colorbarvalue.R'))
}))

plotdatasum<-parLapply(cl, 1:78, function(x){
  modobj<-gammodelsum[[x]]
  plotdata<- plotdata_generate(modobj, "Age", agevector)
  return(plotdata)
})
saveRDS(plotdatasum, paste0(interfileFolder, "/plotdatasum_scale_TRUE_SA12.rds"))
plotdatasum.df<-as.data.frame(matrix(NA, nrow = 1, ncol=17))
names(plotdatasum.df)<-c(names(plotdatasum[[2]])[1:13],"SC_label",  "SCrank", 
                         "PartialRsq", "meanderiv2")
#### SA12 index & SC rank
Matrix12<-matrix(NA, nrow=12, ncol=12)
indexup12 <- upper.tri(Matrix12)
indexsave12 <- !indexup12
Matrix12.index<-Matrix12
#13*12/2=78
Matrix12.index[indexsave12]<-c(1:78)
#S-A connectional axis rank
Matrix12.SCrank<-Matrix12
for (x in 1:12){
  for (y in 1:12){
    Matrix12.SCrank[x,y]<-x^2+y^2
  }
}
Matrix12.SCrank[indexup12]<-NA
Matrix12.SCrank[indexsave12]<-rank(Matrix12.SCrank[indexsave12], ties.method = "average")
# rbind plotdata
for (i in 1:78){
  tmp<-plotdatasum[[i]][,-14]
  tmp$SC_label<-names(plotdatasum[[i]])[14]
  tmp$SCrank<-Matrix12.SCrank[indexsave12][i]
  tmp$PartialRsq<-gamresultsum.SAorder.delLM$partialRsq[i]
  tmp$meanderiv2<-gamresultsum.SAorder.delLM$meanderv2[i]
  plotdatasum.df<-rbind(plotdatasum.df, tmp)
}
plotdatasum.df<-plotdatasum.df[-1, ]

```

## Plots

### 2 Developmental trajectories.

```{r plot_trajectory1}
lmthr <- max(abs(gamresultsum.SAorder.delLM$partialRsq))
ggplot()+
  geom_line(data=plotdatasum.df, aes(x=Age, y=fit.ratio, group=SC_label, color=PartialRsq), size=0.8, alpha=0.8)+
  scale_color_distiller(type="seq", palette = "RdBu",direction = -1, limit=c(-lmthr,lmthr))+
  labs(x="Age (years)", y="SC strength (ratio)")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22, color="black"),aspect.ratio = 0.9,
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        plot.title = element_text(size=15, hjust = 0.5), legend.position = "none")

ggsave(paste0(FigureFolder, '/CV',CVthr, '/SA12_sumSCinvnode_fit/devcurve_Rsq_fit.ratio.tiff'), width=20, height =14, units = "cm")
ggsave(paste0(FigureFolder, '/CV',CVthr, '/SA12_sumSCinvnode_fit/devcurve_Rsq_fit.ratio.svg'), dpi=600, width=15, height =15, units = "cm")
```


```{r plot_trajectory2}
gamresultsum.SAorder.delLM <- gamresultsum.SAorder.delLM %>% mutate(
  meanderv2_c = case_when(
    meanderv2 > mean(meanderv2)+3*sd(meanderv2) ~ NA,
    meanderv2 < mean(meanderv2)-3*sd(meanderv2) ~ NA,
    .default = meanderv2
  )
)
lmthr <- max(abs(gamresultsum.SAorder.delLM$meanderv2_c), na.rm = T)
SC_label_derv2_order <- gamresultsum.SAorder.delLM$parcel[order(gamresultsum.SAorder.delLM$meanderv2)]
plotdatasum.df$SC_label2 <- factor(plotdatasum.df$SC_label, levels=SC_label_derv2_order)
plotdatasum.df <- plotdatasum.df %>% left_join(select(gamresultsum.SAorder.delLM, meanderv2_c, parcel), join_by(SC_label==parcel))
ggplot()+
  geom_line(data=plotdatasum.df, aes(x=Age, y=fit.Z, group=SC_label2, color=meanderv2_c), size=0.8, alpha=0.8)+
  #paletteer::scale_color_paletteer_c("pals::ocean.matter", direction = -1, values=colorbarvalues.meanderiv2, oob = squish) +
  scale_color_distiller(type="seq", palette = "RdBu",limit=c(-lmthr,lmthr), na.value = "#053061", direction = -1)+
  labs(x="Age (years)", y="SC strength (z-score)")+
  #scale_color_manual(values = rev(brewer.pal(10, "RdBu")))+
  scale_y_continuous(breaks = c(-1.5, 0.0, 1.5))+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22, color="black"),aspect.ratio = 0.9,
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        plot.title = element_text(size=15, hjust = 0.5), legend.position = "none")

ggsave(paste0(FigureFolder, '/CV',CVthr, '/SA12_sumSCinvnode_fit/devcurve_meanderv2_fit.Z.tiff'), width=20, height =14, units = "cm")
ggsave(paste0(FigureFolder, '/CV',CVthr, '/SA12_sumSCinvnode_fit/devcurve_meanderv2_fit.Z.svg'), dpi=600, width=15, height =15, units = "cm")

```


## 3. Average fitted values for 10 deciles of connectional axis

```{r average_trajectory}
SA12_10 <- data.frame(SCrank=Matrix12.SCrank[indexsave12])
SA12_10 <- SA12_10 %>%
  mutate(decile = ntile(SCrank, 10))
#table(SA12_10$decile)
SA12_10$SC_label <- paste0("SC.", c(1:78), "_h")
write.csv(SA12_10, paste0(interfileFolder, '/SA12_10.csv'), row.names = F)
plotdatasum.df.label <- merge(plotdatasum.df, SA12_10, by="SC_label", all.x=T)
#names(plotdatasum.df.label)
plotdatasum.df.decile <- plotdatasum.df.label %>%
  group_by(decile, Age) %>%
  summarise(fit.avg = mean(fit), SCranktype_order=mean(decile))

#names(plotdatasum.df.decile)
plotdatasum.df.decile <- plotdatasum.df.decile %>%
  group_by(decile) %>%
  mutate(fit.Z = scale(fit.avg), fit.ratio=fit.avg/fit.avg[1])

ggplot(data=plotdatasum.df.decile, aes(x=Age, y=fit.Z, group=decile, color=decile))+
  geom_line(size=1.5, alpha=0.8)+
  #paletteer::scale_color_paletteer_c("pals::ocean.matter", direction = -1, values=colorbarvalues.meanderiv2, oob = squish) +
  scale_color_distiller(type="seq", palette = "RdBu", direction = -1)+
  labs(x="Age (years)", y="SC strength (z-score)")+
  #scale_color_manual(values = rev(brewer.pal(6, "RdBu")))+
  theme_classic()+
  theme(axis.text=element_text(size=17.2, color="black"), 
        axis.title =element_text(size=17.2, color="black"),aspect.ratio = 0.8,
        plot.background=element_rect(fill="transparent"),legend.position = "none",
        panel.background=element_rect(fill="transparent"),
        plot.title = element_text(size=15, hjust = 0.5))
ggsave(paste0(FigureFolder,'/CV',CVthr,  '/SA12_decile_sumSCinvnode_fit/devcurve_SCrank_fit.Z_SCtype10.tiff'), dpi=600, width=20, height =14, units = "cm")
ggsave(paste0(FigureFolder,'/CV',CVthr,  '/SA12_decile_sumSCinvnode_fit/devcurve_SCrank_fit.Z_SCtype10.svg'), dpi=600, width=15, height =13, units = "cm")


```


