---
title: "S2nd_compositescorePlot_scatterplot"
author: "Xiaoyu xu"
date: "4/17/2024"
output: html_document
---

This script generate the schema of fluid cognition structure and scatter plots between SC strength and fluid cognition for 3 exemplified connections. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(mgcv)
library(parallel)
library(corrplot)
library(ggcorrplot)
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
library(geomtextpath)
interfileFolder <- 'D:/xuxiaoyu/DMRI_network_development/SC_development/interdataFolder_ABCD'
FigureFolder<-'D:/xuxiaoyu/DMRI_network_development/SC_development/Figure_ABCD_final/SA12/CV75/cognition'
resultFolder<-'D:/xuxiaoyu/DMRI_network_development/SC_development/results_ABCD'
functionFolder<-'D:/xuxiaoyu/DMRI_network_development/SC_development/Rcode_SCdevelopment/gamfunction'
source(paste0(functionFolder, "/gamcog.R"))
## load data
SCdata.all<-readRDS(paste0(interfileFolder, '/SCdata_SA12_CV75_sumSCinvnode.sum.msmtcsd.combatNTBfluid.rds'))
SCdata <- SCdata.all
SA12_10 <- read.csv(paste0(interfileFolder, '/SA12_10.csv'))
SA12_10 <- SA12_10[order(SA12_10$SCrank),]

```

## 1.fluid score component

```{r ComponentPlot}
# flanker, DCCS, PSM, lswm, pattern
fluidplot <- data.frame(
  category = c("Inhibition", "Flexibility", "Episodic memory", "Working memory", "Pattern comparison"),
  value = c(10, 10, 10, 10, 10)
)
fluidplot$category <- as.factor(fluidplot$category)
fluidplot$x <- rep(0.5,5)
fluidplot$angle = c(0, 108,  216, -216, -108)
fluidplot$angle[2:5] = fluidplot$angle[2:5]+180
ggplot(fluidplot, aes(x = x, y = value, fill = category)) +
  geom_bar(stat = "identity", width = 1, show.legend=F) +
  coord_polar(theta = "y") +scale_fill_manual(values=brewer.pal(5, "Set3"))+
  xlim(-1, 1)+geom_text(aes(label = category, angle=angle), position = position_stack(vjust = 0.5),size=7)+
  annotate("text", x=-1, y=0, label="Fluid Cognition",size=7)+
  theme_void()

ggsave(paste0(FigureFolder, '/NTB_fluidscore_component.svg'), width = 16, height=16, units = "cm")

```

## 2. cognition ~ SC strength scatter plot
Residuals regressed out age, sex and head motions were used in these plots. Three connections located at different positions along the connectional axis were selected for exemples.

```{r scatterplots}
Cogvar <- "nihtbx_fluidcomp_uncorrected"
CVthr <- 75
SC_Cog_results.df <- readRDS(paste0(interfileFolder, "/SC_Cog_results_", Cogvar,"_CV", CVthr,"_cog2SC.rds"))
# plot data
nonna_index<-which(!is.na(SCdata[ ,Cogvar]))
SCdata.cog<-SCdata[nonna_index,]
SCdata.cog <- SCdata.cog[SCdata.cog$eventname=="baseline_year_1_arm_1",]
SCdata.cog[,str_detect(names(SCdata.cog), "SC.")] <- lapply(SCdata.cog[,str_detect(names(SCdata.cog), "SC.")], as.numeric)
dataname<-"SCdata.cog"
smooth_var<-"age"
covariates<-"sex+mean_fd"
knots<-3
corrmethod<-"pearson"
cl <- makeCluster(8)
clusterExport(cl, varlist = ls(), envir = .GlobalEnv)
invisible(clusterEvalQ(cl, {
  library(mgcv)
  library(parallel)
  library(psych)
  library(gratia)
  library(tidyverse)
  library(reshape)
  source(paste0(functionFolder, "/gaminteraction.R"))
  source(paste0(functionFolder, "/gamcog.R"))
}))
plotdata <- parLapply(cl, 1:78, function(x) {
  SClabel <- grep("SC.", names(SCdata.cog), value=T)[x]
  region <- SClabel
  gamresult.df <-
    gam.fit.cognition(
      region,
      dataname,
      Cogvar,
      smooth_var, 
      covariates, 
      knots,
      set_fx = TRUE, stats_only = FALSE)
  gamresult.df <- as.data.frame(gamresult.df[[2]])
  return(gamresult.df)
})

SC_Cog_results.df$SClabelorder <- c(1:78)
SC_Cog_results.df <- merge(SC_Cog_results.df, SA12_10, by.x="parcel", by.y="SC_label")
SC_Cog_results.df <- SC_Cog_results.df[order(SC_Cog_results.df$SClabelorder),]

N=57
plotdata_N <- plotdata[[N]]
SCrank = SC_Cog_results.df$SCrank[N]
print(paste("The figure below is for connection with a rank of", SCrank))
ggplot(data = plotdata_N)+
  geom_point(aes(x=SCres, y=cogres), color="grey", size=0.8)+
  geom_smooth(aes(x=SCres, y=cogres),linewidth=1.4, method="lm", color="black")+
  labs(x="SC strength", y="Average performance")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22),aspect.ratio = 0.9,
        axis.line = element_line(linewidth = 0.5),axis.ticks = element_line(linewidth = 0.5),
        plot.title = element_text(size=20, hjust = 0.5, vjust=0),
        plot.subtitle = element_text(size=15, hjust = 0.9, vjust=-6),
        plot.margin = margin(t=10, r=10, b=10, l=10, unit = "pt"),
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        legend.position = "none")

ggsave(paste0(FigureFolder, "/", Cogvar, "/SC", N, "_scatterplot.svg"), width=13.5, height=13.5, units = "cm")
ggsave(paste0(FigureFolder, "/", Cogvar, "/SC", N, "_scatterplot.tiff"), width=13.5, height=13.5, units = "cm")


# corr r distribution
SC_Cog_results.df.sig <- SC_Cog_results.df %>% filter(anova.cov.p.fdr < 0.05)
ggplot(data = SC_Cog_results.df.sig, aes(correstimate, y = ..count..)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "gray", position = position_dodge()) +
  labs(x = expression("Correlation ("*italic("r")*")"), y = "Frequency", title = "Cognitive association (N = 72)") +
  scale_x_continuous(breaks = c(-0.11, -0.08, -0.05)) +
  #scale_y_continuous(limits = c(0, 70), breaks = c(0,10, 20,30, 40,50,60,70)) +
  #geom_hline(aes(yintercept = 10), colour = "red", linetype="dashed")+
  theme_classic()+
  theme(axis.text=element_text(size=22, color="black"), 
        axis.title =element_text(size=22),aspect.ratio = 0.9,
        axis.line = element_line(linewidth = 0.5),axis.ticks = element_line(linewidth = 0.5),
        plot.title = element_text(size=20, hjust = 0.5, vjust=0),
        plot.subtitle = element_text(size=15, hjust = 0.9, vjust=-6),
        plot.margin = margin(t=10, r=10, b=10, l=10, unit = "pt"),
        plot.background=element_rect(fill="transparent"),
        panel.background=element_rect(fill="transparent"),
        legend.position = "none")
ggsave(paste0(FigureFolder, "/", Cogvar, "/SigCorrestimateDistribution.svg"), width=13.5, height=13.5, units = "cm")
ggsave(paste0(FigureFolder, "/", Cogvar, "/SigCorrestimateDistribution.tiff"), width=13.5, height=13.5, units = "cm")



```

